---
title: "IS608_HW1"
author: "Daina Bouquin"
date: "February 9, 2016"
output: html_document
---

The following questions pertain to data on the 5,000 fastest growing companies in the US, as 
compiled by Inc. magazine. That data is available [here](https://github.com/jlaurito/CUNY_IS608/blob/master/lecture1/data/inc5000_data.csv)

```{r}
download.file("https://raw.githubusercontent.com/jlaurito/CUNY_IS608/master/lecture1/data/inc5000_data.csv", "inc5000_data.csv", method="curl")
inc5000 <- read.csv("inc5000_data.csv", na.strings = "NA")
# check the header and structure
head(inc5000)
str(inc5000)
```

1. Create a graph that shows the distribution of companies in the dataset by State (i.e. how many are in each state). There are a lot of States, so consider which axis you should use assuming I am using a ‘portrait’ oriented screen
```{r fig.width=8.5, fig.height=11}
library(ggplot2)
library(dplyr)
library(RColorBrewer)

c <- ggplot(inc5000, aes(factor(State))) + geom_bar(fill="#009E73")
c <- c + coord_flip()
c <- c + theme(text = element_text(size=12), axis.title=element_text(size=14,face="bold")) 
c <- c + labs(title = "Number of Companies by State", x= "State", y= "Count")
c <- c + theme(plot.title = element_text(size=22))
c

ggsave('q1.png', height = 11, width = 8.5)
```

2. For the State with the 3rd most companies, create a plot of average employment by industry for companies in this state (only use cases with full data. Your graph should show how variable the ranges are, and exclude outliers. 
```{r fig.width=8.5, fig.height=9}
counts <- as.data.frame(table(inc5000$State))
colnames(counts) <- c("State", "Count")
head(counts) # check structure
sort(counts$Count, TRUE)[3] # kind of hacky I know...
filter(counts, Count == 311)
# pull out only NY data from orinal dataset and remove incomplete cases
ny_data <- filter(inc5000, State == "NY")
ny_data <- ny_data[complete.cases(ny_data),]
str(ny_data)

d <- ggplot(ny_data) +  geom_bar(aes(Industry, Employees, fill = Industry), position = "dodge", stat = "summary", fun.y = "mean", fill="sky blue")
d <- d + coord_flip()
d <- d + theme(legend.position="none")
d <- d + theme(text = element_text(size=12), axis.title=element_text(size=14,face="bold")) 
d <- d + labs(title = "Average Number of Employees by Industry in NY", x= "Industry", y= "Average Number of Employees")
d <- d + theme(plot.title = element_text(size=17))
d

ggsave('q2.png', height = 8.5, width = 9)
```

3. Generate a chart showing which industries generate the most revenue per employee. 
```{r fig.width=8.5, fig.height=9}
#filter out incomplete cases
rev_data <- inc5000[complete.cases(inc5000),]
# Create a new column showing revenue/employee
rev_data <- rev_data %>% mutate(rev_per_em = Revenue / Employees)
str(rev_data)
head(rev_data)

# Plot average revenue/employee by industry
e <- ggplot(rev_data) +  geom_bar(aes(Industry, rev_per_em, fill = Industry), position = "dodge", stat = "summary", fun.y = "mean", fill="darkslateblue")
e <- e + coord_flip()
e <- e + theme(legend.position="none")
e <- e + theme(text = element_text(size=12), axis.title=element_text(size=14,face="bold")) 
e <- e + labs(title = "Average Revenue per Employees by Industry", x= "Industry", y= "Average Revenue per Employees")
e <- e + theme(plot.title = element_text(size=17))
e

ggsave('q3.png', height = 8.5, width = 9)
```





